Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-01-16 paper/1_Initial/manuscript.qmd
fix conflicts @AnafNeves

@@

<!-- what most recent paper? -->

For instance, while the use of subjective self-reports questionnaires to measure deeply embodied functions might seem paradoxical, recent redefinitions of interoception, emphasizing the role of high-level and metacognitive elaboration of interoceptive information, has provided the theoretical grounding to support the idea that some facets of interoception (including participants' metacognitive beliefs) can be assessed subjectively, providing useful and interesting measures [@murphy2019] [**TODO: also cite the most recent paper OF MURPHY with the chinese**]. 
**TODO: that's a long sentence consider splitting**
The notion that self-reports might not reflect the same processes as other interoception measures is important to contextualize the apparent lack of convergence between various measures in the field [@desmedt2022measures]. 
A better understanding of what is being measured with different questionnaires and dimensions, as well their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception.

@@

While the assessment of the validity of an interoception scale can be conceived as theoretically challenging, several traits have been used to assess convergent validity for the the IAS, including expected negative associations with alexithymia [@brand2023; @koike2023; @campos2021, @murphy2019; @lin2023], somatic symptoms [@brand2023; @koike2023], depressive symptoms [@brand2023; @koike2023], anxiety [@brand2023], neuroticism [@brand2023]. 
<!-- **anything else?**. TO BE COMPLETED -- to be determined -->

@@

@@

@@

@@

@@

@@

@@

@@

@@

These findings contrast with previous research, which all found that a 2-factor model, or a 1-factor model fits the data best [@murphy2019, @lin2023, @campos2021, @brand2023].
While this analysis also revealed an okay fit for the 1-factor model, the 4-factor model revealed the best fit.
The 4-factor structure reveals different ‘hubs' of items that are related, not only in this structure analysis, but also in underlying mechanisms.
The ‘wind-burp-cough-sneeze-vomit' category, for example, only entails items that are linked to excretion through the mouth.
The other categories are organized similarly.
This organization and structure is useful for further analysis, as the data can be analyzed and interpreted according to a grouping that is coherent in result, as well as underlying mechanisms.

There are several items that show redundancy suggesting that adapting the IAS would be beneficial for validity.
Based on the given results, we suggest removing the tickle, while keeping the itch item (stats).
Other items with slight redundancy were Hungry and Thirsty, Urinate and Defecate, and Sneeze and Cough.

Interestingly, @lin2023 also found that tickle and itch were redundant, excluding one of them but the reason being that the character for both words is the same in the Chinese language.
On top of that, they came up with a shortened version of the IAS, excluding further items, resulting in a 12-item IAS, which aligns with our findings, suggesting that further items are ambiguous as to whether they should be removed.
In contrast, other findings also found itch and tickle to be redundant but did not suggest excluding items .

@@

The questionnaires used for the IAS correlates are listed in Table 2 (**TODO: add the rest of the questionnaires, sample items and references**).

**Table 2.** Questionnaires used for the IAS correlates analysis.

| Questionnaire              | Number of Dimensions | Assessment                              | Number of Items | Scoring                | Example Item |
|------------|------------|------------|------------|------------|------------|
| **Interoceptive Related**  |                      |                                         |                 |                        |              |
| MAIA-2                     | 8                    | Interoception                           | 37              | 6-point Likert scale   |              |
| BPQ                        | 2                    | Body awareness and autonomic reactivity | 49              | 5-point Likert scale   |              |
| TAS-20                     | 3                    | Alexithymia                             | 20              | 5-point Likert scale   |              |
| BVAQ                       | 2                    | Alexithymia                             | 40              | 5-point Likert scale   |              |
| **Mood**                   |                      |                                         |                 |                        |              |
| BDI-II                     | 1                    | Severity of depressive symptoms         | 21              | 0-to-3-point values    |              |
| PHQ-4                      | 2                    | Anxiety and depressive symptoms         | 4               | 5-point Likert scale   |              |
| STAI-T                     | 1                    | Trait anxiety                           | 20              | 4-point Likert scale   |              |
| GAD-2                      | 1                    | General Anxiety                         | 2               | 4-point Likert scale   |              |
| **Personality**            |                      |                                         |                 |                        |              |
| NEO-FFI                    | 1                    | Neuroticism                             | 12              | 5-point Likert scale   |              |
| Mini IPIP6                 | 6                    | Personality                             | 24              | Analogue scales        |              |
| BFI                        | 5                    | Personality                             | 44              | 5-point Likert scale   |              |
| PID-5-SF                   | 5                    | Dysfunctional personality traits        | 25              | 4-point Likert scale   |              |
| **Psychopathology**        |                      |                                         |                 |                        |              |
| SPQ-BRU                    | 4                    | Schizotypy                              | 32              | 5-point Likert scale   |              |
| MSI-BPD                    | 1                    | Borderline personality disorder         | 10              | Analogue scales        |              |
| ASQ - Short                | 5                    | Autistic Traits                         | 28              | 4-point Likert scale   |              |
| **Beliefs and Misbeliefs** |                      |                                         |                 |                        |              |
| GCB                        | 5                    | Conspiracy beliefs                      | 15              | 5-point Likert scale   |              |
| PI-18                      | 1                    | Beliefs about the world                 | 99              | 6-point Likert scale   |              |
| LIE scale                  | 4                    | Lying tendencies                        | 16              | Visual analogue scales |              |

*Note.* MAIA-2 Multidimensional Assessment of Interoceptive Awareness Version-2 (Mehling et al. 2018); BPQ Body Perception Questionnaire; TAS-20 Toronto Alexithymia Scale (Bagby et al., 1994); BVAQ Bermond-Vorst Alexithymia Questionnaire (Vorst & Bermond, 2001); BDI-II Beck's Depression Inventory (Dozois et al., 1998); PHQ-4 Patient Health Questionnaire (Kroenke et al., 2009); STAI-T State-Trait-Anxiety Inventory Trait-Version (Spielberger et al. 1970); GAD- 2 Generalized Anxiety Disorder Scale – 2 (Spitzer et al., 2007); NEO-FFI NEO Five-Factor Inventory Neuroticism (Costa & McCrae, 2008); Mini IPIP6 Mini International Personality Item Pool (Sibley et al. 2011); BFI Big Five Inventory (John et al., 2010); PID-5-SF Personality Inventory for DSM-5 Short Form (Thimm et al., 2016); SPQ-BRU Schizotypal Personality Questionnaire (Davidson et al., 2016); MSI-BPD Mclean Screening Instrument for Borderline Personality Disorder (Zanarini et al. 2003); ASQ Short Autism-Spectrum Quotient (Hoekstra et al. 2011); GCB Generic Conspiracist Beliefs scale (Brotherton et al., 2013); PI-18 Primals Inventory (Clifton et al., 2021); LIE scale Lying Profile Questionnaire (Makowski, et al,. 2023).

>>>>

[**TO DO**: add - previous work suggests the importance of physiological contexts (Vlemincx et al., 2021)]
**I  would rather put that in the discussion in the suggestions for better scales**

@@

For instance, while the use of subjective self-reports questionnaires to measure deeply embodied functions might seem paradoxical, recent redefinitions of interoception, emphasizing the role of high-level and metacognitive elaboration of interoceptive information, has provided the theoretical grounding to support the idea that some facets of interoception (including participants' metacognitive beliefs) can be assessed subjectively, providing useful and interesting measures [@murphy2019; @lin2023].
**TODO: that's a long sentence consider splitting**
The notion that self-reports might not reflect the same processes as other interoception measures is important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures]. 
A better understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness), is thus needed to clarify the role of self-reports in the assessment of interoception.

@@

While the assessment of the validity of an interoception scale can be conceived as theoretically challenging, several traits have been used to assess convergent validity for the the IAS, including expected negative associations with alexithymia [@brand2023; @koike2023; @campos2021, @murphy2019; @lin2023], somatic symptoms [@brand2023; @koike2023, @lin2023], depressive symptoms [@brand2023; @koike2023; @lin2023], anxiety [@brand2023], neuroticism [@brand2023] and self-esteem [@murphy2019]. 
<!-- **anything else?**. TO BE COMPLETED -- to be determined -->

@@

Few studies have examined the correlation between objective interoceptive measures, such as the Heartbeat Counting Task (HCT; Schandry, 1981) and the Heartbeat Detection Task (HDT; Kleckner et al., 2015). Existing findings report weak correlations for the HCT (Murphy et al., 2019), no correlations except for the sensitivity variable of the HCT (Brand et al., 2023), and small correlations for the HDT (Brand et al., 2023). These results suggest that the IAS, a subjective measure, does not strongly align with objective interoceptive assessments.

<!-- current study -->

@@

<!-- Distibutions of items -->

@@

<!-- Univariate analysis -->

@@

<!-- Network analysis -->

@@

<!-- Stability analysis -->

@@

<!-- Factor analysis -->

@@

<!-- Confirmatory factor analysis -->

@@

<!-- Ias structure -->

@@

<!-- linking findings to previous research -->

These findings contrast with previous research, which all found that 2-factor model[@murphy2019; @koike2023], 1-factor model [@brand2023] and bifactor model [@lin2023; @campos2021] fits the data best.
While this analysis also revealed an okay fit for the 1-factor model, the 4-factor model revealed the best fit.
The 4-factor structure reveals different ‘hubs' of items that are related, not only in this structure analysis, but also in underlying mechanisms.
The ‘wind-burp-cough-sneeze-vomit' category, for example, only entails items that are linked to excretion through the mouth.
The other categories are organized similarly.
This organization and structure is useful for further analysis, as the data can be analyzed and interpreted according to a grouping that is coherent in result, as well as underlying mechanisms.

<!-- Items redundancy -->

There are several items that show redundancy suggesting that adapting the IAS would be beneficial for validity.
Based on the given results, we suggest removing the tickle, while keeping the itch item [**todo**: stats?].
Other items with slight redundancy were Hungry and Thirsty, Urinate and Defecate, and Sneeze and Cough.

<!-- Lins short IAs kept: Hunger, Breath, Urinate, Taste, Vomit, Cough, Temperature, Sexual Arousal, Wind, Muscle, Pain, Itch -->

Interestingly, @lin2023 also found that tickle and itch were redundant, excluding one of them but the reason being that the character for both words is the same in the Chinese language.
On top of that, they came up with a shortened version of the IAS, excluding further items, resulting in a 12-item IAS, which aligns with our findings, suggesting that further items are ambiguous as to whether they should be removed.
In contrast, other findings also found itch and tickle to be redundant but did not suggest excluding items [@campos2021].

<!-- low density on extreme values (1 and 5) -->

@@

The questionnaires used for the IAS correlates are listed in **Table 2** (**TODO: add the rest of the questionnaires, sample items and references**).

\input{../../analysis/figures/table2.tex}
